---
title: "Explore the Online News Popularity Dataset"
output: pdf_document
---

```{r setup, include=FALSE}
source("src/load.R")
```

### See plans document for general information about the dataset.

```{r str}
str(news)
```

This is how the distribution of the number of shares looks like:

```{r shares_distribution}
quantile(news$shares, seq(0,1,0.1))
quantile(news$shares, c(0, 0.001, 0.005, 0.01, 0.05, seq(0.1,0.9,0.1), 0.95, 0.99, 0.995, 0.999))
```

From this we can see that the distribution is extremely long tailed therefore we shall attempt to predict the *log(shares)* value instead of the plain shares.

### Number of articles published on each day of the week: 

```{r weekday distribution}
weekday_data <- melt(news[,c(weekday_columns), with = F], measure.vars = weekday_columns)
weekday_data <- weekday_data[,"value":= sum(value), by = variable]
weekday_data <- unique(weekday_data)
weekday_data[,"variable":=factor(variable,
                                 levels = c("monday", "tuesday", "wednesday",
                                            "thursday", "friday", "saturday", "sunday"))]
ggplot(weekday_data, aes(x = variable, y = value)) + 
    geom_bar(fill = "darkblue", stat = "identity")
```

Nothing special or unexpected here.

###Number of articles published in each data channel:

```{r data channel distribution}
data_channel_data <- melt(news[,c(data_channel_columns), with = F],
                          measure.vars = data_channel_columns)
data_channel_data <- data_channel_data[,"value":= sum(value), by = variable]
data_channel_data <- unique(data_channel_data)
ggplot(data_channel_data, aes(x = variable, y = value)) + 
    geom_bar(fill = "darkblue", stat = "identity")
```

### Number of articles not assigned to any channel:
```{r without channel}
wo_channel <- replace_channel_columns(news)
nrow(wo_channel[channel == "no_channel",])
```

### Articles per week:  
```{r timedelta}
ggplot(news, aes(x = timedelta)) + geom_histogram(binwidth = 7) + 
    xlab("days since article publishing to dataset publishing")
```

The same with aggregation to 30 days:

```{r timedelta per month}
ggplot(news, aes(x = timedelta)) + geom_histogram(binwidth = 30) + 
    xlab("days since article publishing to dataset publishing")
```

```{r summary}
summary(news)
```

###More about the distribution of the shares variable

```{r shares}
ggplot(news[shares<10000,], aes(x = shares)) + geom_density()
ggplot(news, aes(x = log(shares))) + geom_density()
```

The following was a discovery but it does not effect the core of the predictive modeling.

```{r mod}
ggplot(news[shares < 1000,], aes(x = shares%%10)) + geom_histogram(binwidth = 1)
ggplot(news[shares > 1000,], aes(x = shares%%10)) + geom_histogram(binwidth = 1)
```


###Distribution of number of shares by data channel:

```{r shares_by_channel}
channel_data <- news[,c(data_channel_columns, "shares"),with = F]
channel_data <- channel_data[,c("lifestyle", "entertainment", "bus", "socmed", "tech", "world") :=
                                 list(lifestyle * shares, entertainment * shares,
                                      bus * shares, socmed * shares, tech * shares, world * shares)]
channel_data <- melt(channel_data[,data_channel_columns, with = F],
                     measure.vars = data_channel_columns)
channel_data <- channel_data[value != 0]
ggplot(channel_data, aes(x = variable, y = value)) + geom_boxplot()
channel_data <- channel_data[,c("mean_shares", "median_shares"):=
                                 list(mean(value), median(value)), by = variable]
channel_data <- unique(channel_data[,.(variable, mean_shares, median_shares)])
channel_data <- melt(channel_data,measure.vars = c("mean_shares", "median_shares"),
                     id.vars = "variable" )
ggplot(channel_data, aes(x = variable)) + 
    geom_bar(aes(y = value, fill = `variable.1`), stat = "identity", position = "dodge")
```

###Distribution of number of shares by weekday:

```{r shares_by_weekday}
weekday_data <- news[,c(weekday_columns, "shares"),with = F]
weekday_data <- weekday_data[,c("monday", "tuesday", "wednesday",
                                "thursday", "friday", "saturday", "sunday") :=
                                 list(monday * shares, tuesday * shares,
                                      wednesday * shares, thursday * shares,
                                      friday * shares, saturday * shares, sunday * shares)]
weekday_data <- melt(weekday_data[,weekday_columns, with = F],
                     measure.vars = weekday_columns)
weekday_data <- weekday_data[value != 0]
ggplot(weekday_data, aes(x = variable, y = value)) + geom_boxplot()
weekday_data <- weekday_data[,c("mean_shares", "median_shares"):=
                                 list(mean(value), median(value)), by = variable]
weekday_data <- unique(weekday_data[,.(variable, mean_shares, median_shares)])
weekday_data <- melt(weekday_data,measure.vars = c("mean_shares", "median_shares"),
                     id.vars = "variable" )
ggplot(weekday_data, aes(x = variable)) + 
    geom_bar(aes(y = value, fill = `variable.1`), stat = "identity", position = "dodge")
```

###Interdependencies betweek the columns LDA_..

```{r lda_distributions}
lda <- news[sample(.N, 1000),lda_columns, with = F]
plot(lda)
lda <- melt(lda, measure.vars = lda_columns)
ggplot(lda, aes(x = value)) + geom_density(aes(col = variable))
```

###Correlation with target variable:

```{r corr}
options(scipen = 4)

correlations <- sapply(news, function(x) {cor(x, news$shares)})

columns <- as.data.table(names(news))
columns$corr_with_target <- 0
for (name in columns$V1) {
    columns[V1 == name]$corr_with_target <- correlations[[name]]
}

columns <- columns[order(-corr_with_target),]
columns
```
